Storm surges in the Strait of Georgia simulated with a regional model

### 2015-01-05 AO_guides/manuscriptAO.tex
Keith's comment about gamma^2 in December 2006 case

%overview oF SoG/Salish Sea models
The early modelling work of \citet{crean1988mathematical,crean1988numerical} and \citet{stronach1993update} resulted in a series of models covering the Strait of Juan de Fuca, the Strait of Georgia, and Puget Sound using a finite difference discretization of the equations of motion. %A barotropic version of their model was used to reproduce storm surge water level elevations with reasonable accuracy, however, this barotropic version did not include the effects of fresh water entering the system from rivers. Later, a baroclinic version of their model examined residual circulation and fresh water flushing rates \citep{stronach1993update, marinone1996model}.
These models pointed to an important connection between the model's bottom friction parametrization and the location and amplitude of the M$_2$ amphidrome. They found that a local increase in the bottom friction coefficient, typically from 0.002 to 0.004, especially in the area of Haro Strait, was necessary in order to adequately reproduce the observed M$_2$ tidal amplitudes and phase lags in the Strait of Georgia \citep{stronach1993update}. Without local adjustments to the bottom friction their M$_2$ amplitudes were too low and the M$_2$ tidal phases were too large in the Strait of Georgia.  Other studies over a similar domain, such as the barotropic finite element model of \citet{foreman1995tidal}, have avoided local increases in bottom friction. \citet{foreman2004m} used data assimilation techniques to study M$_2$ tidal dissipation without local changes to the bottom friction in a two-dimensional finite element model that did not include baroclinic effects.

@@

Modelled residuals are generally too low in this hindcast (Figure \ref{fig:dec2006}), however, the timing of the surge agrees well. All Willmott skill scores rank above 0.88 and the $\gamma^2$ measure also indicates good agreement between model and observations (Table \ref{tab:statistics}) but the performance has degraded when compared to the Feb 2006 and Nov 2006 cases. This could be due to a number of factors. First, the very large wind speeds observed during this storm are significantly underrepresented by the atmospheric model leading to lower wind stress and less piling of water against the coast. Second, the anomaly forcing at the open boundaries could be too low. The observed residual at Tofino, BC (a point near by but outside of the domain) is up to 10 cm lower than the observed residual at Neah Bay, WA (a point near by but inside of the domain). Since the storm surge is less sensitive to wind forcing at the surface than anomaly forcing at the open boundaries, the latter is likely the cause.

>>>>

%overview oF SoG/Salish Sea models
The early modelling work of \citet{crean1988mathematical}, \citet{crean1988numerical} and \citet{stronach1993update} resulted in a series of models covering the Strait of Juan de Fuca, the Strait of Georgia, and Puget Sound using a finite difference discretization of the equations of motion. %A barotropic version of their model was used to reproduce storm surge water level elevations with reasonable accuracy, however, this barotropic version did not include the effects of fresh water entering the system from rivers. Later, a baroclinic version of their model examined residual circulation and fresh water flushing rates \citep{stronach1993update, marinone1996model}.
These models pointed to an important connection between the model's bottom friction parametrization and the location and amplitude of the M$_2$ amphidrome. They found that a local increase in the bottom friction coefficient, typically from 0.002 to 0.004, especially in the area of Haro Strait, was necessary in order to adequately reproduce the observed M$_2$ tidal amplitudes and phase lags in the Strait of Georgia \citep{stronach1993update}. Without local adjustments to the bottom friction their M$_2$ amplitudes were too low and the M$_2$ tidal phases were too large in the Strait of Georgia.  Other studies over a similar domain, such as the barotropic finite element model of \citet{foreman1995tidal}, have avoided local increases in bottom friction. \citet{foreman2004m} used data assimilation techniques to study M$_2$ tidal dissipation without local changes to the bottom friction in a two-dimensional finite element model that did not include baroclinic effects.

@@

Modelled residuals are generally too low in this hindcast (Figure \ref{fig:dec2006}), however, the timing of the surge agrees well. All Willmott skill scores rank above 0.88 and the $\gamma^2$ measure also indicates good agreement between model and observations (Table \ref{tab:statistics}). The $\gamm^2$ measure suggests that the standard deviation of the error in the modelled resdiuals is roughly half of the standard deviation of the observed residuals. Note that the performance has degraded when compared to the Feb 2006 and Nov 2006 cases. This could be due to a number of factors. First, the very large wind speeds observed during this storm are significantly underrepresented by the atmospheric model leading to lower wind stress and less piling of water against the coast. Second, the anomaly forcing at the open boundaries could be too low. The observed residual at Tofino, BC (a point near by but outside of the domain) is up to 10 cm lower than the observed residual at Neah Bay, WA (a point near by but inside of the domain). Since the storm surge is less sensitive to wind forcing at the surface than anomaly forcing at the open boundaries, the latter is likely the cause.
